Approximating the Optimal State Abstraction

### 2016-04-30 icml_workshop/introduction.tex
adding some more structure to tex

Abstraction plays a fundamental role in learning. Through abstraction, intelligent agents may reason about only the salient features of their environment while ignoring what is irrelevant. Consequently, agents are able to solve considerably more complex problems than they would be able to without the use of abstraction. However, \textit{exact abstractions}, which treat only fully-identical situations as equivalent, require complete knowledge that is often computationally intractable to obtain. Furthermore, often no two situations are identical, so exact abstractions are often ineffective. To overcome these issues, we investigate \textit{approximate abstractions} that enable agents to treat sufficiently similar situations as identical. This work characterizes the impact of equating ``sufficiently similar'' states in the context of planning and \ac{RL} in \acp{MDP}. The remainder of our introduction contextualizes these intuitions in \acp{MDP}.

% Abstraction is a good idea b/c MDP has poly comp comp and samp comp in S, but S grows fast.
Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{papadimitriou1987complexity,littman1995complexity}. Similarly, many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, the size of the state space of an \ac{MDP} grows super-polynomially with the number of variables that characterize the domain, a result of Bellman's curse of dimensionality. Thus, solutions polynomial in state space size are often ineffective for sufficiently complex tasks. For instance, a robot involved in a pick-and-place task might be able to employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider grows exponentially with the number of objects with which it is working \cite{abel2015goal}.

% Existing work on state abstraction.
Thus, a key research agenda for planning and \ac{RL} is leveraging abstraction to reduce large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. This agenda has given rise to methods that reduce \textit{ground} \acp{MDP} with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states according to some notion of equality or similarity. In the context of \acp{MDP}, we understand exact abstractions as those that aggregate states with equal values of particular quantities, for example, optimal $Q$-values. Existing work has characterized how exact abstractions can fully maintain optimality in \acp{MDP}~\cite{li2006towards,dean1997modelmin}.

>>>>

\dnote{Abstraction pitch}
% Introduce a pitch about abstraction that's a bit different from our other paper.
Intelligent agents that leverage abstraction may reason about just the salient features of their environment, while ignoring what is irrelevant. Consequently, agents are able to solve considerably more complex problems than they would be able to without the use of abstraction. However, \textit{exact abstractions}, which treat only fully-identical situations as equivalent, require complete knowledge that is often computationally intractable to obtain. Furthermore, often no two situations are identical, so exact abstractions are often ineffective. To overcome these issues, we investigate \textit{approximate abstractions} that enable agents to treat sufficiently similar situations as identical. This work characterizes the impact of equating ``sufficiently similar'' states in the context of planning and \ac{RL} in \acp{MDP}. The remainder of our introduction contextualizes these intuitions in \acp{MDP}.

% Abstraction is a good idea b/c MDP has poly comp comp and samp comp in S, but S grows fast.
%Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{papadimitriou1987complexity,littman1995complexity}. Similarly, many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, the size of the state space of an \ac{MDP} grows super-polynomially with the number of variables that characterize the domain, a result of Bellman's curse of dimensionality. Thus, solutions polynomial in state space size are often ineffective for sufficiently complex tasks. For instance, a robot involved in a pick-and-place task might be able to employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider grows exponentially with the number of objects with which it is working \cite{abel2015goal}.

\dnote{Prior work summary}
% Existing work on state abstraction.
A key research agenda for planning and \ac{RL} is leveraging abstraction to reduce large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. This agenda has given rise to various methods of abstraction that reduce the size of \acp{MDP} with large state spaces; \dnote{Overview of abstraction methods}

\dnote{Close with a lack of clarity about what is meaningful to retain from an abstraction theory}

In this work, we propose two desiderata on theories of abstraction in the context of planning or learning in an \ac{MDP}; the first we call \textsc{Compressibility}, which stipulates that a theory on abstraction ought to achieve {\it compression}, while the second we coin \textsc{Optimality}, stating that abstractions ought to maintain at least some notion of optimality. We defend these desiderata briefly in the next section, and describe the status of current work relative to these two desiderata. In light of these desiderata, we propose a formal framework that allows for simple state-abstraction methods as well as extension to temporal abstractions. We summarize the current state of the art theoretical results under this framework and describe how this framework satisfies both desiderata. Moreover, a central goal of this work is to illuminate future work on options and temporally extended actions. We argue that the present state of the literature is inconclusive as to the effectiveness of options in general, and that, the theory of temporal abstraction afforded by the state aggregation approach allows for the satisfaction of both \textsc{Compressibility} and \textsc{Optimality}. We offer a description of this framework, as well as preliminary empirical results, and paths to relevant analysis.

% ---- OLD STUFF ----
In the context of \acp{MDP}, we understand exact abstractions as those that aggregate states with equal values of particular quantities, for example, optimal $Q$-values. Existing work has characterized how exact abstractions can fully maintain optimality in \acp{MDP}~\cite{li2006towards,dean1997modelmin}.
